---
editor_options: 
  markdown: 
    wrap: 72
---

[[**VECTORS**]{.underline}]{style="RED"}

Vectors come in two types:- (i)Atomic (ii)Lists Atomic - all elements
must have same type -

**Atomic Vectors**- (i)Logical (ii)Integer --- Character (iii)double ---
Vectors (iv)Character

Logical-T/F Doubles-(0.1234) Integers-(1234L,1e4L) Strings-" ",' '

**Test**- is.\*() function - which type it is. for e.g-
is.logical(),is.integer() etc.

**Coerced**- When we attempt to combine different types - coerced in a
fixed order: Character -\> Double -\> Integer -\> Logical.

for e.g - Combining a chr and int -\> chr.

**(or)**

you can coerce by using as.\*() function. For e.g - as.logical(),
as.double() etc.

**Attributes**------

```{r}
a <- 1:3
attr(a,"x") <- "abcdef"
attr(a,"y") <- 4:6
str(attributes(a))
```

```{r}
a <- structure(
  1:3,
  x = "abcdef",
  y = 4:6
)
str(attributes(a))
```

**Names**-- 3 ways to do it -- (i)When creating it:

```{r}
x <- c(a=1,b=2,c=3)
```

(ii)By assigning a character vector to names()

```{r}
x<-1:3
names(x)<-c("a","b","c")
```

(iii)Inline, with setNames:

```{r}
x<- setNames(1:3,c("a","b","c"))
```

**Dimensions**--

```{r}
a<-matrix(1:6,nrow=2,ncol=3)
b<-array(1:12,c(2,3,2))
c<-1:6
dim(c)<-c(3,2)
```

**Vector, Matrix, & Array:**----

```{r}
#1d Vector
str(1:3)
#column vector
str(matrix(1:3,ncol=1))
#row vector
str(matrix(1:3,nrow = 1))
#array vector
str(array(1:3,3))
```

**S3 Atomic Vectors:**--

Class-attribute turns an object into an **S3 object**.

**Categories**--

(i)Factor--Categorical Data(Integer)

(ii)Date--Dates with day resolution(Double)

(iii)POSIXct--date-times with second or sub-second resolution(Double)

(iv)Difftime--Durations

**FACTORS:**

It is of 2 attributes : a class and a factor

```{r}
x<-factor(c("a","b","b","a"))
x

typeof(x)
attributes(x)
```

```{r}
s_char<-c("m","m","m")
s_factor<- factor(s_char,levels = c("m","f"))
table(s_char)
table(s_factor)
```

*Ordered*factors - low,medium and High levelling

```{r}
grade<-ordered(c("b","b","a","c"),levels=c("a","b","c"))
grade
```

**DATES:**

Double vectors

Have only class "Date" and no other attributes

```{r}
today<-Sys.Date()
today
typeof(today)
attributes(today)
```

```{r}
date<-as.Date("2022-08-12")
unclass(date)#The no. of days since 1970-01-01
#Time started for Unix computers
```

**DATE-TIMES:(POSIX:** Portable Operating System Interface**)**

It provides two ways of storing date-time information,

*POSIXct(*calendar time*) & POSIXlt(*local time*)*

```{r}
now_ct<-as.POSIXct("2022-08-12 3:00",tz="UTC")
now_ct
typeof(now_ct)
attributes(now_ct)
```

```{r}
structure(now_ct, tzone="Asia/Tokyo")
structure(now_ct, tzone="America/New_York")
structure(now_ct, tzone="Australia/Lord_Howe")
structure(now_ct, tzone="Europe/Paris")
```

**DURATIONS:**

The amount of time between pairs of Dates

```{r}
one_week<-as.difftime(1,units="weeks")#units can be days,weeks
one_week
typeof(one_week)
attributes(one_week)
```

**LISTS:**

It can be of any type, not just vectors.

(i) **Creating -**

```{r}
l1 <- list(
  1:3,
  "a",
  c(TRUE,TRUE,TRUE),
  c(2.3,5,9)
)
typeof(l1)
str(l1)
```

**Recursive** vectors - Lists - as it can contain another list.

```{r}
l2<-list(list(list(1)))
str(l2)

l3<-list(list(1,2),c(3,4))
l4<-c(list(1,2),c(3,4))#c combines several lists into one
str(l3)
str(l4)
```

(iii) **Matrices and Arrays -**

```{r}
l<-list(1:3,"a",TRUE,1.0)
dim(l)<-c(2,2)
l
l[[1,1]]
```

**DATA FRAMES AND TIBBLES:**

First, vector then to list and List is converted into two types called

(i) data.frame & (ii) tibble

**DATA FRAME -** It is a named list of vectors with attributes for
(column) names, row.names, and its class, "data.frame":

```{r}
df1<-data.frame(x=1:3,y=letters[1:3])
df1
typeof(df1)
attributes(df1)
```

The length of each vectors should be same \<- Rectangle \<- share
properties of both Matrices and lists.

(i) It has `rownames()`and `colnames().`The `names()` of a data frame
are the column names.

(ii) \-\-\-\-\-\-\-\-\-- `nrow()`rows and`ncol()`columns.`length()`:
gives the number of columns.

**TIBBLES -(***Lazy and Surly: They do less and complain more***)**

Replacement for Data frames.

```{r}
library(tibble)

df2<-tibble(x=1:3,y=letters[1:3])
typeof(df2)
attributes(df2)
```

(i) **Creating -**

Same as Data Frame

*LAZY* :- As it doesn't coerce the inputs

```{r}
data.frame(x=1:4,y=1:2)

tibble(x=1:4,y=1)
```

# 
